---
title: "Worldclouds - Music Lyrics"
author: "Lester Pi"
date: "2/25/2017"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
header-includes: \usepackage[dvipsnames]{xcolor}
---

```{r Setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
library(base)
library(tm)
library(SnowballC)
library(wordcloud)
library(sqldf)
library(knitr)
library(rmarkdown)
library(tcltk)

```

```{r global_options, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(fig.path='Figs/')
```

```{r set Width, echo=FALSE, warning=FALSE, message=FALSE}
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
opts_chunk$set(dev = 'pdf')
```


#Introduction

Wordclouds are cool, and what better way to play around with them than music lyrics?

The data I am working with includes 380,000+ lyrics from MetroLyrics spanning 10 different genres (omitted "NA" and "Other" genres). Let's see what we can find.

Being a fan of Rap and Hip-Hop, what better way to start things off with than...


```{r,echo=FALSE, warning=FALSE, message=FALSE}
setwd("C:/cygwin64/home/Lester/wordclouds")
lyrics = read.csv("lyrics.csv",stringsAsFactors = FALSE)
genres = sqldf("select genre from lyrics group by genre")

#split into genres
country = sqldf("select * from lyrics where genre='Country'")
electronic = sqldf("select * from lyrics where genre='Electronic'")
folk = sqldf("select * from lyrics where genre='Folk'")
hiphop = sqldf("select * from lyrics where genre='Hip-Hop'")
indie = sqldf("select * from lyrics where genre='Indie'")
jazz = sqldf("select * from lyrics where genre='Jazz'")
metal = sqldf("select * from lyrics where genre='Metal'")
pop = sqldf("select * from lyrics where genre='Pop'")
rb = sqldf("select * from lyrics where genre='R&B'")
rock = sqldf("select * from lyrics where genre='Rock'")
```



```{r,echo=FALSE, warning=FALSE, message=FALSE}
#add n word as stopword for obvious reasons
#remove punc
stopWords = c(removePunctuation(stopwords('SMART')),"nigga","niggas","niggaz")
#data - dataset
#pal - color palette
#val - value of color palette
#stem - 1 for stemming on
buildWordCloud = function(data, pal,val,stem, name){
  corpus = Corpus(VectorSource(data$lyrics))
  corpus = tm_map(corpus, content_transformer(tolower))
  corpus = tm_map(corpus, PlainTextDocument)
  corpus = tm_map(corpus, removePunctuation)
  corpus = tm_map(corpus, removeWords, stopWords)
  if(stem==1) corpus = tm_map(corpus, stemDocument)
  wordcloud(corpus, max.words = 75, random.order = FALSE, colors=brewer.pal(val,pal), main = name)
}

```

#The Evolution of Hip Hop

By subsetting the data by decades, it will be easy to see how word usage has changed over time. For obvious reasons, I added some non-traditional stopwords to the analysis. Also take note that the pre 90s Hip-Hop subset has less than 150 observations, while the other eras have plenty more.

The wordclouds on the left have stemming off and the ones on the right have stemming on.


```{r,echo=FALSE, warning=FALSE, message=FALSE}
#subset data by years
hiphop80s = sqldf("select * from hiphop where (year > 1900 and year < 1990)")
hiphop90s = sqldf("select * from hiphop where (year >= 1990 and year < 2000)")
hiphop00s = sqldf("select * from hiphop where (year >= 2000 and year < 2010)")
hiphop10s = sqldf("select * from hiphop where (year >= 2010 and year < 2020)")
```

\begin{center}

\textbf{Pre 1990s Hip Hop}
```{r, fig.align="center",echo=FALSE, warning=FALSE, message=FALSE}
par(mfrow=c(1,2))
buildWordCloud(hiphop80s,"Spectral",8,0, "Pre 90s Hip Hop")
buildWordCloud(hiphop80s,"Spectral",8,1, "Pre 90s Hip Hop")
```


\textbf{1990s Hip Hop}
```{r, fig.align="center",echo=FALSE, warning=FALSE, message=FALSE}
par(mfrow=c(1,2))
buildWordCloud(hiphop90s,"Set1",8,0, "1990-2000")
buildWordCloud(hiphop90s,"Set1",8,1, "1990-2000")
```



\textbf{2000s Hip Hop}
```{r, fig.align="center",echo=FALSE, warning=FALSE, message=FALSE}
par(mfrow=c(1,2))
buildWordCloud(hiphop00s,"Dark2",8,0, "2000 - 2010")
buildWordCloud(hiphop00s,"Dark2",8,1, "2000 - 2010")
```



\textbf{2010s Hip Hop}
```{r, fig.align="center",echo=FALSE, warning=FALSE, message=FALSE}
par(mfrow=c(1,2))
buildWordCloud(hiphop10s,"Paired",8,0, "2010 and Onward")
buildWordCloud(hiphop10s,"Paired",8,1, "2010 and Onward")
```


\end{center}


It's pretty easy to see Hip-Hop has gotten a lot more vulgar since its conception.

#But Wait, There's More!

Just for fun, here's worldclouds, with stemming on, of other genres!

\begin{center}
Top Row: \textcolor{BlueViolet}{Pop}, \textcolor{BlueViolet}{Ja}\textcolor{ForestGreen}{zz}, \textcolor{ForestGreen}{R\&B}

Middle Row: \textcolor{BlueViolet}{Elec}\textcolor{ForestGreen}{tro}\textcolor{Dandelion}{nic}, \textcolor{darkgray}{Metal}, \textcolor{Dandelion}{Cou}\textcolor{BurntOrange}{nt}\textcolor{red}{ry}

Bottom Row: \textcolor{BurntOrange}{Rock}, \textcolor{BurntOrange}{Fo}\textcolor{red}{lk}, \textcolor{red}{Indie}
\end{center}

```{r,fig.align="center",echo=FALSE, warning=FALSE, message=FALSE}
par(mfrow=c(3,3))
buildWordCloud(pop,"Blues",9,1, "Pop")
buildWordCloud(jazz,"BuGn",9,1, "Jazz")
buildWordCloud(rb,"Greens",9,1, "R&B")
buildWordCloud(electronic,"YlGnBu",9,1, "Electronic")
buildWordCloud(metal,"Greys",9,1, "Metal")
buildWordCloud(country,"YlOrRd",9,1, "Country")
buildWordCloud(rock,"Oranges",9,1, "Rock")
buildWordCloud(folk,"OrRd",9,1, "Folk")
buildWordCloud(indie,"Reds",9,1, "Indie")
```

#Closing Comments

Personally, I find it pretty funny most generes of music have "love" as their top word, besides Metal and Hip-Hop. Does this mean most music was derived from love? Or could it be a sign that music was created as a mating ritual  (assuming mating biologically evolved into our concept of love)?

What does this say about the genres of music not as focused on love? I belive it shows the nature of metal and hip hop. I don't listen to metal music, so I can't comment without genearlizing the whole genre. However, I can say that not all Hip-Hop is about misogyny and violence, and at least in my view, the quality Hip-Hop has subject matter deeper than soley materialistic matters. At least for me, this just goes to show that "good" Hip-Hop is hard to find, especially in recent times.

Wordclouds are pretty fun!

#Sources

Kaggle: https://www.kaggle.com/gyani95/380000-lyrics-from-metrolyrics

MetroLyrics: Original source of Kaggle data

